---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# geocreep

<!-- badges: start -->
[![R-CMD-check](https://github.com/tobiste/geocreep/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/tobiste/geocreep/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of geocreep is to quick and easy calculate strain rates, water fugacities and deviatoric stress from grain sizes using R.
The codes use Monte Carlo simulations to propagate uncertainites of the flow parameters into the final estimate.

## Installation

You can install the development version of geocreep from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("tobiste/geocreep")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(geocreep)

# Set seed for reproducibility
set.seed(20250411)
```

### Water fugacity

Calculate water fugacity from temperature and pressures using the Pitzer and Sterner (1994) equation:
```{r fug}
temperature <- units::set_units(300, degC)
pressure <- units::set_units(400, MPa)

fugacity <- ps_fugacity(pressure, temperature)
print(fugacity)
```
### Grain-size piezometry
Calculating deviatoric stress from grain size (using the Stipp and Tullis perometer):

```{r gs}
grainsize <- units::set_units(11, um)
stress <- grainsize_piezometry(grainsize, model = "Stipp-reg2-3")

# Median
print(stress$median)

# 68% Interpercentile range
print(stress$ir_68)
```

### Flow laws
Calculate strain rate using a defined flow law from stress, temperature, and fugacity: 

```{r edot}
edot <- strain_rate(stress = stress$median, temperature = temperature, fugacity = fugacity, model = "Hirth2001")

# Median
print(edot$median)

# 68% Interpercentile range
print(edot$ir_68)
```
