---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# geocreep

<!-- badges: start -->
[![R-CMD-check](https://github.com/tobiste/geocreep/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/tobiste/geocreep/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of geocreep is to quick and easy calculate strain rates, fugacities and differential stress from grain sizes using R.
The codes use Monte Carlo simulations to propagate uncertainties of the flow parameters into the final estimate.

## Installation

You can install the development version of geocreep from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("tobiste/geocreep")
```

## Example

```{r example, message=FALSE, warning=FALSE}
library(geocreep)
library(units) # load this package to define your units

# Set seed for reproducibility
set.seed(20250411)
```


### Water fugacity

Calculate water fugacity from temperature and pressures using the Pitzer and Sterner (1994) equation:
```{r fug}
nmc <- 1e3 # create 1,000 samples for temperature and pressure using a normal distribution

#Define temperature and pressure
temperature <- units::set_units(rnorm(nmc, 300, 50/1.96), degC)
pressure <- units::set_units(rnorm(nmc, 400, 100/1.96), MPa)

# Calculate fugacity
fugacity <- ps_fugacity(pressure, temperature)

fugacity_stats <- summary(fugacity)
```

> Here, the Monte Carlo simulation produces 1,000 normal-distributed estimates for the fugacity.


### Grain-size piezometry
Calculating differential stress from grain size (e.g. using the Stipp and Tullis, 2003, piezometer):

```{r gs}
# Define grain size
grainsize <- units::set_units(11, um)

# Calculate equivalent differential stress
stress <- grainsize_piezometry(grainsize, model = "Stipp-reg2-3")

# Summary stats of Monte Carlo simulation
summary(stress)
```

>By default, the functions create 1,000,000 samples for each parameter with uncertainties, creating 1e6 results. 
Results of Monte Carlo simulations based on flow law equations do not follow a normal (symmetric) distribution because some parameters are exponents of the power laws. 
Therefore, the median and interpercentile range provide the best estimators for average and dispersion, rather than mean and standard deviation.

Note: There is also a subgrain-size piezometer (Goddard et al. 2021): `subgrainsize_piezometer()`


### Flow laws
Calculate strain rates using a defined flow law for quartz from differential stress, temperature, and fugacity: 

```{r edot}
# Calculate strain rates using temperature, fugacity, and differential stress
edot <- creep_quartz(
  stress = stress, 
  temperature = units::set_units(rnorm(1e6, 300, 50/1.96), degC),
  fugacity = rnorm(1e6, fugacity_stats$mean, fugacity_stats$sd), 
  model = "Hirth2001"
  )

# Summary stats of Monte Carlo simulation
summary(edot)
```

## Author

Tobias Stephan ([tstephan\@lakeheadu.ca](mailto:tstephan@lakeheadu.ca){.email})

## Feedback, issues, and contributions

I welcome feedback, suggestions, issues, and contributions! If you have found a bug, please file [here](https://github.com/tobiste/geocreep/issues) with minimal code to reproduce the issue.

## License
GPL-3.0 License
