---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# geocreep

<!-- badges: start -->
[![R-CMD-check](https://github.com/tobiste/geocreep/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/tobiste/geocreep/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of geocreep is to quick and easy calculate strain rates, fugacities and differential stress from grain sizes using R.
The codes use Monte Carlo simulations to propagate uncertainties of the flow parameters into the final estimate.

## Installation

You can install the development version of geocreep from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("tobiste/geocreep")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example, message=FALSE, warning=FALSE}
library(geocreep)
library(units) # load this package to define your units

# Set seed for reproducibility
set.seed(20250411)
```


### Water fugacity

Calculate water fugacity from temperature and pressures using the Pitzer and Sterner (1994) equation:
```{r fug}
#Define temperature and pressure
temperature <- units::set_units(300, degC)
pressure <- units::set_units(400, MPa)

# Calculate fugacity
fugacity <- ps_fugacity(pressure, temperature)

print(fugacity)
```
### Grain-size piezometry
Calculating differential stress from grain size (e.g. using the Stipp and Tullis piezometer):

```{r gs}
# Define grain size
grainsize <- units::set_units(11, um)

# Calculate equivalent differential stress
stress <- grainsize_piezometry(grainsize, model = "Stipp-reg2-3")

# Median differential stress
print(stress$median)

# 68% Interpercentile range of MC estimates
print(stress$ir_68)
```

Note: There is also a subgrain-size piezometer (Goddard et al. 2021): `subgrainsize_piezometer()`


### Flow laws
Calculate strain rates using a defined flow law for quartz from differential stress, temperature, and fugacity: 

```{r edot}
# Calculate strain rates using temperature, the fugacity, and the MC estimates for differential stress calculated before
edot <- creep_quartz(stress = stress$samples, temperature = temperature, fugacity = fugacity, model = "Hirth2001")

# Median strain rate
print(edot$median)

# 68% Interpercentile range of MC estimates
print(edot$ir_68)
```

## Author

Tobias Stephan ([tstephan\@lakeheadu.ca](mailto:tstephan@lakeheadu.ca){.email})

## Feedback, issues, and contributions

I welcome feedback, suggestions, issues, and contributions! If you have found a bug, please file [here](https://github.com/tobiste/geocreep/issues) with minimal code to reproduce the issue.

## License
GPL-3.0 License
